Sperm should evolve to make female meiosis fair.

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
made the tec compile

%FIgures 
{\bf Abstract:}
Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
Because only one of four products of female meiosis is transmitted to
offspring, o\"{o}genesis is inherently susceptible to such conflicts,
as alleles can compete to be transmitted into the egg. 
%(meiotic drive).  
Genomics conflicts drive the rapid evolution of suppression mechanisms, and
	many facets of  o\"{o}genesis and gametogenesis have been 
	interpreted as mechanisms of protection against genomic outlaws. 
Why then is female meiosis often left uncompleted until after fertilization - 
	potentially providing an opportunity for sperm alleles to
        meddle with its outcome in their own favor? %and facilitate meiotic drive? 
The population genetic theory presented herein suggests that
	sperm nearly always evolve to increase the fairness of female
        meiosis in the face of genomic conflicts. %%felt this
                                %%might be an important addition as
                                %%they will spread if no cost to organism
These results are consistent with current knowledge of sperm-dependent meiotic drivers 
(loci whose distortion of female meiosis depends on sperm genotype),}
	and suggest that the requirement of fertilization for the
        completion of female meiosis, and variation in stage of meiosis that
        fertilization occurs at may represent a mechanism employed by females 
	 to ensure a fair meiosis. %prevent meiotic drive.
\newpage

@@

%%%sperm-acting drive suppressors
While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
\gc{Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, i.e. in repulsion phase, creating a
        third allele $B^{-}$, Model 4. \gc{This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       is does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes.} This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     \ref{Eggsperm_3_allele_cartoon}, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and eqns \ref{A-} and \ref{unlinked}).  
This allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive.

>>>>

%FIgures 
{\bf Abstract:}
Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
Because only one of four products of female meiosis is transmitted to
offspring, o\"{o}genesis is inherently susceptible to such conflicts,
as alleles can compete to be transmitted into the egg. 
%(meiotic drive).  
Genomics conflicts drive the rapid evolution of suppression mechanisms, and
	many facets of  o\"{o}genesis and gametogenesis have been 
	interpreted as mechanisms of protection against genomic outlaws. 
Why then is female meiosis often left uncompleted until after fertilization - 
	potentially providing an opportunity for sperm alleles to
        meddle with its outcome in their own favor? %and facilitate meiotic drive? 
The population genetic theory presented herein suggests that
	sperm nearly always evolve to increase the fairness of female
        meiosis in the face of genomic conflicts. %%felt this
                                %%might be an important addition as
                                %%they will spread if no cost to organism
These results are consistent with current knowledge of sperm-dependent meiotic drivers 
(loci whose distortion of female meiosis depends on sperm genotype),
	and suggest that the requirement of fertilization for the
        completion of female meiosis, and variation in stage of meiosis that
        fertilization occurs at may represent a mechanism employed by females 
	 to ensure a fair meiosis. %prevent meiotic drive.
\newpage

@@

%%%sperm-acting drive suppressors
While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, i.e. in repulsion phase, creating a
        third allele $B^{-}$, Model 4. \gc{This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       is does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes.} This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     \ref{Eggsperm_3_allele_cartoon}, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and eqns \ref{A-} and \ref{unlinked}).  
This allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive.
